Spatial visualization

### 2014-03-05 article2013.Rmd
päivitin slaidit iltapäivän ja moskovan esiyksiin

library(texreg)
texreg(list(soc.4,ind.4,socVsInc),
       #float.pos="H",
       #use.packages=TRUE,
       #booktabs=TRUE,
       #dcolumn=TRUE,
       scriptsize=TRUE,
       naive = TRUE,
       caption.above=TRUE,
       doctype=FALSE,
       custom.model.names = model.names.multi.2,
       custom.coef.names = coef.names.multi,
       include.pvalues=TRUE,
       caption="Logistic multilevel random intercept model comparing the full models for both dependent variable with model predicting social blame against individual blame")

@@

To sum up, the results indicate that in general social blame type of explanation is clearly the most popular explanation in post-socialist countries. Social blame is followed by individual blame, social fate and individual fate types of explanations. Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty in the post-socialist space. In regard to differences between the East and the West of Europe, results show that Western Europeans emphasise more social reasons for poverty.

As assumed results also emphasised large between-country differences among post-socialist countries. By employing a multilevel approach we wanted to shed light especially on this issue. To begin with, we found that differences between countries are responsible for almost 30 percent of the variation in the support for social blame and 12 percent for individual blame type of explanation. Our contextual level predictors decreased the variance down to 15 percent for social and 4 percent for individual type of explanation.

As we hypothesised changes in country’s economic performance during the period of global financial crises were associated with attributions for poverty. In fact, changes in economic performance was the most significant predictor of our country-level variables for both social and individual blame types of explanations: the better the economy had performed the more support gained the individual blame type of explanation. And on the other hand, increased financial strain is associated with stronger support for the view that poverty is caused injustices of society.

@@

Results of the individual-level determinants of perceptions supported the self-interest hypothesis which states that those who have financial difficulties perceive the causes of poverty differently than those who are not exposed to poverty-related risks. Findings indicate that those whose main source of income rely mainly on social or private transfers, those who perceive their income level to be lower than country’s median, those who perceive their future economic situation to be worse than now as well as those who perceive that the global financial crises has affected negatively to their household’s economic situation endorse more likely external (social blame) than internal (individual blame) reasons for poverty. Thus, analyses showed that that it would be fruitful to include a perspective of risk groups when studying attitudes toward poverty, and also to include future expectations in the equation (also Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).

This study focused only on poverty-related risks as individual-level determinants. There are, of course, other variables which might be associated with attributions even more than risk factors. For instance, another line of research have emphasised the role of other beliefs, attitudes and values rather than interest (e.g. Cozzarelli et al., 2001). Treating the values, attitudes and other perceptions as determinants for attributions of poverty is, however, problematic because it is difficult to prove the causal relationship between attributions and other attitudes. This does not mean that they are irrelevant. Conversely, we argue that future research should try to investigate to what extent high levels of social and individual blame are associated with wider welfare attitudes in post-socialist countries. In order, for example, to explore policy implications more thoroughly, there is a need for future research to focus on the relationship between attributions for poverty and the legitimacy of specific types of welfare policies in post-socialist countries.

As a methodological contribution, findings showed interesting differences between the East and the West regarding the effects of contextual variables. Even though the selection of Western European countries in this study included countries from different welfare regimes, the differences between countries in economic change, democracy and income inequality are quite modest compared especially to CIS countries. Consequently our results showed more modest and in some cases opposite associations between dependent and contextual variables among Western European countries than among post-socialist group.

@@

@@

>>>>

@@

To sum up, the results indicate that in general social blame type of explanation is clearly the most popular explanation in post-socialist countries. Social blame is followed by individual blame, social fate and individual fate types of explanations. Hence, the results are in line with our assumptions and prior research (Habibov, 2011; Lepianka et al., 2010) regarding the general support for different explanations of poverty in the post-socialist space. In regard to differences between the East and the West of Europe, results show that Western Europeans emphasize more social reasons for poverty.

As assumed results also emphasized large between-country differences among post-socialist countries. By employing a multilevel approach we wanted to shed light especially on this issue. To begin with, we found that differences between countries are responsible for almost 30 percent of the variation in the support for social blame and 12 percent for individual blame type of explanation. Our contextual level predictors decreased the variance down to 15 percent for social and 4 percent for individual type of explanation.

As we hypothesized changes in country’s economic performance during the period of global financial crisis were associated with attributions for poverty. In fact, changes in economic performance was the most significant predictor of our country-level variables for both social and individual blame types of explanations: the better the economy had performed the more support gained the individual blame type of explanation. And on the other hand, increased financial strain is associated with stronger support for the view that poverty is caused by injustices of society.

@@

Results of the individual-level determinants of perceptions supported the self-interest hypothesis which states that those who have financial difficulties perceive the causes of poverty differently than those who are not exposed to poverty-related risks. Findings indicate that those whose main source of income rely mainly on social or private transfers, those who perceive their income level to be lower than country’s median, those who perceive their future economic situation to be worse than now as well as those who perceive that the global financial crisis has affected negatively to their household’s economic situation endorse more likely external (social blame) than internal (individual blame) reasons for poverty. Thus, analyses showed that that it would be fruitful to include a perspective of risk groups when studying attitudes toward poverty, and also to include future expectations in the equation (also Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).

This study focused only on poverty-related risks as individual-level determinants. There are, of course, other variables which might be associated with attributions even more than risk factors. For instance, another line of research have emphasized the role of other beliefs, attitudes and values rather than interest (e.g. Cozzarelli et al., 2001). Treating the values, attitudes and other perceptions as determinants for attributions of poverty is, however, problematic because it is difficult to prove the causal relationship between attributions and other attitudes. This does not mean that they are irrelevant. Conversely, we argue that future research should try to investigate to what extent high levels of social and individual blame are associated with wider welfare attitudes in post-socialist countries. In order, for example, to explore policy implications more thoroughly, there is a need for future research to focus on the relationship between attributions for poverty and the legitimacy of specific types of welfare policies in post-socialist countries.

As a methodological contribution, findings showed interesting differences between the East and the West regarding the effects of contextual variables. Even though the selection of Western European countries in this study included countries from different welfare regimes, the differences between countries in economic change, democracy and living conditions are quite modest compared especially to CIS countries. Consequently our results showed more modest and in some cases opposite associations between dependent and contextual variables among Western European countries than among post-socialist group.

@@

Habibov, Nazim N. (2011) Public Beliefs Regarding the Causes of Poverty during Transition. Evidence from the Caucasus, Central Asia, Russia, and Ukraine. International Journal of Sociology and Social Policy 31 (1-2), 53-74.

@@

Appendix
=======================================================

\setcounter{table}{0}
\renewcommand\thetable{\arabic{table}}

library(texreg)
texreg(list(soc.4,ind.4,socVsInc),
       float.pos="H",
       #use.packages=TRUE,
       #booktabs=TRUE,
       #dcolumn=TRUE,
       scriptsize=TRUE

### 2014-03-05 finalPlots.Rmd
päivitin slaidit iltapäivän ja moskovan esiyksiin

library(RColorBrewer)
library(mapproj)
library(maptools)
library(rgdal)
library(grid)
map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 
                                layer="lits_all_simple"))

@@

ppi <- 300
png("figure/mapplot.png", width=14/2.54*ppi, height=12/2.54*ppi, res=ppi)
mapplot
dev.off()

>>>>

library(RColorBrewer)
library(mapproj)
library(maptools)
library(rgdal)
library(grid)
map <- suppressWarnings(readOGR(dsn="/home/aurelius/data/shapefiles/lits/", 
                                layer="lits_all_simple"))

@@

ppi <- 300
png("figure/mapplot.png", width=15/2.54*ppi, height=9/2.54*ppi, res=ppi)
mapplot
dev.off()

### 2014-03-05 mapPlots.Rmd
päivitin slaidit iltapäivän ja moskovan esiyksiin

```{rmapPlotsLoadData}
library(RColorBrewer)
library(mapproj)
library(maptools)
library(rgdal)
load("~/workspace/lits/attributions/data/lits.RData")
map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 
                                layer="lits_all_simple"))

@@

```{rmapPlotsGroup1}
ggplot(map.df, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
  scale_fill_manual(values=cbPalette) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
  coord_map(project="orthographic") +
  theme(legend.position="top")

@@

```{rmapPlotsDV1}
### Faceted showing population shares of explanations
library(scales)
plotMap <- merge(map.df,macro,by.x="NAME",by.y="cntry")
plotMap <- plotMap[order(plotMap$order),]

@@

### Perceived Gini coefficient

```{rmapPlotsIvMacro2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = perGini)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("blue"), high=muted("red"), midpoint=33.55) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

>>>>

```{rmapPlotsLoadData}
library(RColorBrewer)
library(mapproj)
library(maptools)
library(rgdal)
load("~/workspace/lits/attributions/data/lits.RData")
map <- suppressWarnings(readOGR(dsn="/home/aurelius/data/shapefiles/lits/", 
                                layer="lits_all_simple"))

@@

```{rmapPlotsGroup1, fig.height=13, fig.width=16}
ggplot(map.df, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
  scale_fill_manual(values=cbPalette) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
  coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,60)) +
  theme(legend.position="top")

@@

```{rmapPlotsDV1}
### Faceted showing population shares of explanations
library(scales)
plotMap <- merge(map.df,macro,by.x="NAME",by.y="cntry")
plotMap <- plotMap[plotMap$year == 2010,]
plotMap <- plotMap[order(plotMap$order),]

@@

### Human Development Index

```{rmapPlotsIvMacro2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = undp_hdi)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("red"), high=muted("blue"), midpoint=0.7570) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```
